Scipion Tailored for TEM Facilities: Image Processing in Streaming

### 2018-02-20 0.abstract.tex
w

Three dimensional electron microscopy (3DEM) is becoming a very data-intensive field in which vast amounts of experimental data are acquired at high speed. To manage such large-scale image data  projects, we have previously described  a modular workflow system called \scipion. We present here the first main extension of \scipion that allows to process data in streaming. Streaming processing allows to overlap data acquisition with the first steps of digital image processing and therefore helps to: (1) detect problems at early stages, (2) save computing time and (3) provide users with a in deep evaluation of the data quality before the data acquisition is finished. This idea of processing the data while being acquired has been implemented in different laboratories mainly using custom-made scripts. The advantages of our \scipion solution are: (1) flexibility: users can customize their workflows from a  large collection of preloaded algorithms; (2) repeatability and traceability: workflows are stored and can be re-executed later, either with the same data or a different data set; (3) monitorization: there are special utilities that constantly check how the execution of the algorithms is going and inform about possible anomalies; (4) reports: we have developed several GUIs that are updated periodically and produce a graphical summary (e.g, CTF defocus values, system load, etc). This summary may be  generated in HTML format that can be easily published though a public web-site to provide access for external users; (5) backup: both local and remote backup of the data and image processing is provided. At present, \scipion has been installed and is in production mode in three facilities. And it is being considered in several more.

>>>>

Three dimensional electron microscopy (3DEM) is becoming a very data-intensive field in which vast amounts of experimental data are acquired at high speed. To manage such large-scale image data  projects, we have previously described  a modular workflow system called \scipion \citep{conesa2018}. We present here the first main extension of \scipion that allows to process data in streaming. Streaming processing allows to overlap data acquisition with the first steps of digital image processing and therefore helps to: (1) detect problems at early stages, (2) save computing time and (3) provide users with a in deep evaluation of the data quality before the data acquisition is finished. This idea of processing the data while being acquired has been implemented in different laboratories mainly using custom-made scripts. The advantages of our \scipion solution are: (1) flexibility: users can customize their workflows from a  large collection of preloaded algorithms; (2) repeatability and traceability: workflows are stored and can be re-executed later, either with the same data or a different data set; (3) monitorization: there are special utilities that constantly check how the execution of the algorithms is going and inform about possible anomalies; (4) reports: we have developed several GUIs that are updated periodically and produce a graphical summary (e.g, CTF defocus values, system load, etc). This summary may be  generated in HTML format that can be easily published though a public web-site to provide access for external users; (5) backup: both local and remote backup of the data and image processing is provided. At present, \scipion has been installed and is in production mode in four Cryo-EM facilities. And it is being considered in many more.

### 2018-02-20 1.introduction.tex
w

Many software packages have been developed to provide automated data collection, such as SerialEM \citep{Mastronarde2005}, Leginon \citep{Suloway2009}, UCSF-Tomography \citep{Zheng2007}, Tom Toolbox \citep{Nickell2005}, EPU \citep{EPU}, Latitude in Digital Micrograph \citep{Latitude}, or EMMenu \citep{emmenu}, etc. However, the next step,  which is to provide image processing while  data are being collected, is not so well served. In fact, in most places this is accomplished through home made scripts. Exceptions are UCSFImage4 \citep{Li2015} and Focus \citep{Biyani2017}. As we will explain here, \scipion differs from these solutions in a larger flexibility both from the point of view of users and facility staff.

>>>>

Many software packages have been developed to provide automated data collection, such as SerialEM \citep{Mastronarde2005}, Leginon \citep{Suloway2009}, UCSF-Tomography \citep{Zheng2007}, Tom Toolbox \citep{Nickell2005}, EPU \citep{EPU}, Latitude in Digital Micrograph \citep{Latitude}, or EMMenu \citep{emmenu}, etc. However, the next step,  which is to provide image processing while  data are being collected, is not so well served. In fact, in most places this is accomplished through home made scripts (i.e. \citep{Pichkur2018}). Exceptions are UCSFImage4 \citep{Li2015} and Focus \citep{Biyani2017}. As we will show in this article, \scipion differs from these solutions in a larger flexibility both from the point of view of users and facility staff.

### 2018-02-20 3.workflow.tex
going home

%Comment change in scale in the three examples: CNB, SNC and DIAMOND
%more local, SNC national, Diamond international. Portal infrastructure has
%different levels.

>>>>

The data flow created by an electron microscope requires a careful design of the IT infrastructures.  Different facilities has adopted different solutions and therefore, 
how \scipion  access data and interchange information is very facility dependent. In this section we discuss  the different solutions applied in the different facilities in which \scipion is running.

### 2018-02-20 3.workflow_cnb.tex

We are going to structure this section in three pats. In the first one we discuss the network setup that allow us to move the data from the microscope to the user laboratory. Then we present the standard image processing workflow and finally we comment on the additional software (different from \scipion) used in process data.

\subsubsection{Networking}

@@

The data flow is a follows: (a) the microscope system stores movies in the microscope-NAS, (b) when executing alignment-movie protocols \scipionbox will temporary read the movies from 
the microscope-NAS but will not store them locally, (c) each hour, output-NAS will connect to \scipionbox and copy or update the last \scipion project, (d) at the same time output-NAS may be either executing a script to incrementally backup the \scipion project and the movies to a USB disk or serve this information through the Internet and (e) in parallel \scipionbox transfers to  output-NAS different html reports that may be checked by users or staff with a HTML browser (see \url{http://nolan.cnb.csic.es/scipionbox}).

Although we try to discourage the use of external USB hard disks versus network. At present, this is the preferred option for more of our users.  USB 3.0 drives can theoretically handle data flows of 300 MB/s, but our experience is that, in this busy environment, many of them  have trouble exceeding 50 MB/s. In the ideal situation both raw and processed data are transfered as soon as available. In this way, access to the servers seldom overlaps between two users and once a microscope shift ends, users do not need to wait for their data to be transferred. fggdfgdsgdsfgdf. Comment on ssh cifers

comment on iperf and smbdata

\subsubsection{EMadmin}

>>>>

This section in divided in two parts. In the first one we discuss the network setup that allow us to move the data from the microscope to the user laboratory. Then  we comment on an additional software (called  \emadmin) that connects \scipion with the microscope acquisition 
software (EPU) and records the facility activity.

\subsubsection{Network Setup}

@@

The data flow is a follows: (a) the microscope system stores movies in the microscope-NAS, (b) when executing alignment-movie protocols \scipionbox will temporary read the movies from 
the microscope-NAS but will not store them locally, (c) each hour, output-NAS will connect to \scipionbox and copy or update the last \scipion project, (d) at the same time output-NAS may be either executing a script to incrementally backup the \scipion project and the movies to a USB disk or serve this information through the Internet and, (e) in parallel, \scipionbox transfers to  output-NAS different html reports that may be checked by users or staff with a HTML browser (see \url{http://nolan.cnb.csic.es/scipionbox}).

Although we try to discourage the use of external USB hard disks versus network as way of retrieving the raw and processing data, this is the preferred option for more of our users.
USB 3.0 drives can theoretically handle data flows of 300 MB/s, but our experience is that, in this busy environment, many of them  have trouble exceeding 50 MB/s. In most cases, 50 MB/s is enough to transfer the data before the following microscope shift starts. Nevertheless, real-world is tough and it happens that two, or even three,  users try to download their data simultaneously. Fortunately, output-NAS works as front-end and isolates the data acquiring and processing infrastructure from the user downloading. Network users access to output-NAS using a chrooted common account. This account only allows rsync connections and make impossible to retrieve data if the name of the project to be downloaded is unknown. In order to speed up transfer weak ciphers (as \emph{arcfour}) has been activated in output-NAS.

\subsubsection{\emadmin}

\emadmin was born as a script that created a normalized tree of directories  in which to store the movies acqured by the microscope and help users to choose between a few predetermined image processing workflows. At present it was grown in a client-server application that in addition to the old script capacities it has incorporate the following functions:

\begin{enumerate}
 \item Transparent addition of new workflows. \scipion is able to export any executed workflow as a json file. \emadmin can import this files as use them to create new predetermined workflows.
 \item Automatic creation of pdf user's report containing: microscope acquisition parameters and histograms summarizing data resolution and  defocus. 
 \item The above mentioned information is stored in a database for further analysis. By default the system plots the  average resolution and astigmatism of the different acquisitions versus time
 \item Remote management. \emadmin is a client server application, the client maybe executed in a machine different from the one that executes \scipion
 \item Automatic backup of the \scipion project
 \item \scipion is able to create live html reports on the image acquisition process. \emadmin keeps track of all of them and facilitates the access to this information
\end{enumerate}

\emadmin is free software available at github repository (\url{https://github.com/rmarabini/webservices/tree/master/EMadmin}). You are welcomed to try it but the software is provided as it is and no support or maintenance service is offered.

### 2018-02-20 5.conclusions.tex
move the workflow section on CNB to a more general place
going home

\section{Discussion}
[*+TODO]: Discuss the obtained numbers in the result section

@@

The new generation of high-end cryo electron microscopes and direct electron detectors are behind the recent increase in cryo-EM  popularity. However, the run and maintenance costs of these new microscopes are high and can only be justified keeping the microscopes up and running 24 hours a day. In this way, microscopes tent to belong to centralize facilities where users are allocated microscope time (shifts) to perform a variety of research. Available shifts are a scarce commodity and therefore users (and staff) need to be able to monitor the quality of the acquisition process and get as soon as possible a first glimpse of the reconstructed specimen in order to guaranty an early detection of any issue.

 Despite workflow systems have been explored and used extensively in bioinformatics, there is little progress in that direction in the cryo-EM community. \scipionbox has been designed following and strict workflow paradigm. This approach is specially useful for tasks that need to go from a given start (movies) to a defined end (3D map) but there are many different paths (algorithms) to get there.

 In a workflow approach, users (including microscope staff) can reasonably be expected to edit the paths. They can potentially play with the pipeline in ways that aren't possible when having a programmer write code for them. Furthermore, no having the processing path hardcoded in the program means you don't have to recompile, so changes can be quickly tested   even at run time.

 \scipionbox workflow approach provides provenance (transparency) since for any created object (micrograph, CTFs, class, etc) it is stored which program creates it, using which parameters, and the exact path followed to create it.  In this way,  real tracking and reproducibility are provided.

\scipionbox is part of the \scipion package which is provided 
freely as open source software. Online documentation
describing \scipion download and installation is available
at \url{http://scipion.cnb.csic.es/m/download_form/} and \url{https://github.com/I2PC/scipion/wiki/How-to-Install}, respectively.\\\\

>>>>

\section{Discussion}
[**TODO]: Discuss the obtained numbers in the result section
[**REmove this section?]

@@

The new generation of high-end cryo electron microscopes and direct electron detectors are behind the recent increase in cryo-EM  popularity. However, the run and maintenance costs of these new microscopes are high and can only be justified by keeping the microscopes up and running 24 hours a day. Therefore, microscopes tend to belong to centralized facilities where users are allocated microscope time (shifts) to perform a variety of tasks. Available shifts are a scarce commodity and therefore users (and staff) need to be able to monitor the quality of the acquisition process and obtain as soon as possible a first glimpse of the reconstructed specimen in order to provide an early detection of any issue.

 Workflow systems have been explored and used extensively in bioinformatics but there is relatively little progress in that direction in the cryo-EM community. \scipion has been designed following a strict workflow paradigm. This approach is specially useful for tasks that need to go from a given start (movies) to a defined end (3D map) but there are many different paths (algorithms) and steps to get there.

 In a workflow approach, users (including microscope staff) can reasonably be expected to edit the paths. They can potentially play with the pipeline in ways that aren't possible when having a programmer write code for them. Furthermore, not having the processing path hardcoded in the program means you don't have to recompile, so changes can be quickly tested   even at run time. The \scipion workflow approach provides provenance, that is,  for any created object (micrograph, CTFs, class, etc) \scipion stores metadata such as which programs, which parameters, and which exact path was followed to create it.  In this way,  real tracking and reproducibility are provided.

 \scipion stream capabilities can handle large and continuously produced volumes of data. The benefits of streaming are clear: (1) it helps to process data as soon as it is generated,  providing almost instant results on data quality; (2) it saves user processing time at their home institutions since part of the processing is done while data is being acquired; and (3) 
 it further reduces the overall experimental time by allowing processing the data while it is transferred to the final user laboratory. \scipion has been deployed in cloud and there is a  a public image, with all the required preinstalled cryoEM software,  available at Amazon WebServices EC2 and EGI federated clouds. These images make use of \scipion streaming capabilities, that is, they are able to process data as soon as it has been transfered to an Amazon storage unit, and can be used to handle  production peaks.

\scipion is provided 
freely as open source software. Online documentation
describing \scipion download and installation is available
at \url{http://scipion.cnb.csic.es/m/download_form/} and \url{https://github.com/I2PC/scipion/wiki/How-to-Install}, respectively. Cloud installation instruction are available at \url{https://github.com/I2PC/scipion/wiki/Scipion-in-the-Cloud}. Other public software mentioned in this article as \emadmin, XXXX and YYYY are available at \url{https://github.com/rmarabini/webservices/tree/master/EMadmin}, www eee respectively.
If you would like more information about how to install \scipion or you have suggestions on how to improve it contact us at scipion@cnb.csic.es.
